# Data quality & cleaning

## Missing values

**Only four variables have more than 30% missing values. Most of them are below 10%**

```{r eval=TRUE, echo=FALSE, warning=FALSE, include=TRUE} 
#Load Libraries: p_load can install, load,  and update packages
pacman::p_load(rstudioapi,dplyr, ggplot2, lubridate, devtools, tidyr,magrittr, 
               lemis,zoo,gdata, knitr, DT, plotly, countrycode,htmltools, 
               highcharter,crosstalk)

# Set working directory 
path <- getwd()
knitr::opts_knit$set(root.dir = normalizePath(path.expand(path), 
                                              winslash = ("/"), 
                                              mustWork = TRUE))

# Reading dataset 
data<-readRDS("../data/data.rds")

# Transform some variables to factor/numeric/datetime
data %<>% mutate_at(c("control_number", "species_code", "taxa", "class", "genus",
                      "species", "subspecies", "specific_name", "generic_name",
                      "description", "country_origin", "country_imp_exp", 
                      "purpose","source", "action", "disposition", 
                      "disposition_year", "shipment_year", "import_export",
                      "port", "us_co", "foreign_co", "unit"), as.factor)

# Remove scientific notation
options(scipen=999)
rm(path)

```

```{r eval=TRUE, echo=TRUE, warning=FALSE, include=TRUE} 
# Are there missing values?
missing_values <- data %>%
  gather(key = "key", value = "val") %>%
  mutate(is_missing = is.na(val)) %>%
  group_by(key, is_missing) %>%
  summarise(num_missing = n()/nrow(data) *100) %>%
  filter(is_missing==T) %>%
  select(-is_missing) %>%
  arrange(desc(num_missing)) 

plot_ly(missing_values, x=~reorder(key, desc(num_missing)), y=~num_missing, color=~key) %>% 
  add_bars() %>%
  add_text(text=~round(num_missing),textposition = 'top') %>%
  layout(title = "<b>Missing values per variable</b>",
         xaxis= list(title= "<b>Variable</b>" ,tickangle=-65),
         yaxis = list(title = "<b>Number of missing values</b>"))

rm(missing_values)
```

### Missing values in "value" year over year
**If we have a look at the "value" variable (dollars), we can notice that there isn't enough information about it in the data produced between 2008-2011 and 2013-2014. Specifically, the years 2008, 2009, 2010 and 2014 only contain missing values**

```{r eval=TRUE, echo=TRUE, warning=FALSE, include=TRUE} 
missing_value<- data %>%
  select(shipment_year, value) %>%
  group_by(shipment_year) %>%
  summarise(missing = sum(is.na(value)/n())*100, 
            complete= round(100 -   sum(is.na(value)/n())*100,2)) %>%
  gather("key", "percentage", missing, complete)

ggplot(data=missing_value, aes(x=shipment_year, y=percentage, fill=key)) +
  geom_bar(position = "dodge", stat = "identity") +
  theme_bw() +  facet_grid(~shipment_year,scales = "free_x") +   scale_fill_manual(values=c("springgreen4", "rosybrown"))

rm(missing_value)
```

## Excluding plants and microorganisms

**We classified 98.93 % of data by kingdoms using the class, taxa and genus columns**. This way,  we excluded those elements belonging to fungi, plantae, bacteria, chromista and unspecified kingdoms. 

* **Bacterias**: Gammaproteobacteria

* **Chromista**: Phaeophyceae

* **Plantae**: Cycadopsida, Liliopsida, Magnoliopsida, Pinopsida, Polypodiopsida

* **Fungi**: Agaricomycetes, Chytridiomycetes	

* **Animalia**: Actinopterygii, Amphibia, Anthozoa, Arachnida, Ascidiacea, 
Asteroidea, Aves, Bivalvia, Branchiopoda, Calcarea, Cephalaspidomorphi,
Cephalopoda, Cestoda, Chilopoda, Clitellata, Cubozoa, Demospongiae, Diplopoda,
Echinoidea, Elasmobranchii, Enteropneusta, Eurotatoria,Gastropoda, Gymnolaemata,
Hexactinellida, Hexanauplia, Holocephali, Holothuroidea, Hoplonemertea, Hydrozoa,
Insecta, Leptocardii, Malacostraca, Mammalia, Maxillopoda,Merostomata, Myxini,
Ophiuroidea, Ostracoda, Pilidiophora, Polychaeta, Polyplacophora, Pycnogonida,
Reptilia, Sarcopterygii, Scaphopoda, Scyphozoa, Secernentea, Sipunculidea,
Tentaculata, Thaliacea, Trematoda

```{r eval=TRUE, echo=TRUE, warning=FALSE, include=TRUE} 
# classifying classes by kingdom 
classes<-data %>% drop_na(class) %>% group_by(class) %>% summarise(N=n())

animalia_kingdom <- c("Actinopterygii", "Amphibia", "Anthozoa", "Arachnida",
                      "Ascidiacea", "Asteroidea", "Aves", "Bivalvia",
                      "Branchiopoda", "Calcarea", "Cephalaspidomorphi",
                      "Cephalopoda", "Cestoda", "Chilopoda", "Clitellata",
                      "Crinoidea", "Cubozoa", "Demospongiae", "Diplopoda",
                      "Echinoidea", "Elasmobranchii", "Enteropneusta",
                      "Eurotatoria", "Gastropoda", "Gymnolaemata",
                      "Hexactinellida","Hexanauplia", "Holocephali",
                      "Holothuroidea", "Hoplonemertea", "Hydrozoa", "Insecta",
                      "Leptocardii", "Malacostraca", "Mammalia", "Maxillopoda",
                      "Merostomata", "Myxini", "Ophiuroidea", "Ostracoda",
                      "Pilidiophora", "Polychaeta", "Polyplacophora",
                      "Pycnogonida", "Reptilia", "Sarcopterygii", "Scaphopoda",
                      "Scyphozoa", "Secernentea", "Sipunculidea", "Tentaculata",
                      "Thaliacea", "Trematoda")

fungi_kingdom <-c("Agaricomycetes", "Chytridiomycetes")

plantae_kingdom<- c("Cycadopsida", "Liliopsida", "Magnoliopsida", "Pinopsida", 
                    "Polypodiopsida", "Ulvophyceae")

bacterias_kingdom <- c("Gammaproteobacteria")

chromista_kingdom <- c("Phaeophyceae")
```

```{r eval=TRUE, echo=TRUE, warning=FALSE, include=TRUE}
# Adding kingdom level based on class
data$kingdom<- ifelse(data$class %in% animalia_kingdom, "Animalia",
               ifelse(data$class %in% fungi_kingdom, "Fungi",        
               ifelse(data$class %in% plantae_kingdom, "Plantae", 
               ifelse(data$class %in% bacterias_kingdom, "Bacteria",                 
               ifelse(data$class %in% chromista_kingdom, "Chromista", "Other"                              )))))              # 403,596 unclassified

# Adding kingdom level based on genus
data$kingdom[which(data$genus=="Other live inverts" | data$genus=="Animals" |
                   data$genus=="Corals"| data$genus=="Dugesia"|  
                   data$genus=="Nemertea" | data$genus=="Mollusca" |
                   data$genus=="Xenoturbella" | data$genus=="Chondrichthyes" |
                   data$genus=="Chordata" | data$genus=="Paracatenula" |
                   data$genus=="Porifera")] <- "Animalia"
                                                      # 299,674 unclassified

# Adding kingdom level based on taxa
data$kingdom[which(data$taxa=="crustacean" | data$taxa=="fish" | 
                     data$taxa=="coral"| data$taxa=="shell")] <- "Animalia" 
                                                      #  58,883 unclassified

data$kingdom[which(data$taxa=="plant")] <- "Plantae"  #  58,855 unclassified

# Let's exclude those elements from fungi, plantae, bacteria, chromista and unspecified kingdoms. 
data <- data %>% filter(kingdom=="Animalia") # 5,512,706 --> 5,451,832 rows

rm(classes, animalia_kingdom, bacterias_kingdom, chromista_kingdom,
   plantae_kingdom, fungi_kingdom)

```

```{r eval=TRUE, echo=TRUE, warning=FALSE, include=TRUE}
# Renaming 
data <- data %>% 
  
  # Renaming units
  mutate(unit = recode(unit, "NO" = "Number", "KG" = "Kilograms", 
                       "GM"= "Grams", "M2"="Square_meters", 
                       "LT"= "Liters", "ML"= "Milliliters", 
                       "MT"="Meters", "CM" = "Centimeters", 
                       "C3"= "Cubic_centimeters", "C2"="Square_centimeters",
                       "MG"="Miligrams", "M3"="Cubic_meters"),
         
           # Renaming import_export       
          import_export = recode(import_export, "I" = "Import"),
         
         # Renaming levels
          action = recode(action, "C" = "Cleared", "R"= "Refused"),
         
         # Renaming countries
          country_origin = countrycode(country_origin, "iso2c", 
                                  "country.name", nomatch = NULL),
         
         
          country_origin = recode(country_origin, "AN" = "Netherlands Antilles",
                                "XX" = "Unknown","ZZ" = "High Seas"),
         
          country_imp_exp = countrycode(country_imp_exp, "iso2c", 
                                  "country.name", nomatch = NULL),
  
          country_imp_exp = recode(country_imp_exp, "AN" = "Netherlands Antilles",
                                "XX" = "Unknown","ZZ" = "High Seas"),
         
         # Renaming ports
          port = recode(port, "AG" = "Agana, GU", "AL" = "Alcan, AK", "AN" = "Anchorage, AK",
                        "AT" = "Atlanta, GA", "BA" = "Baltimore, MD", "BL" = "Blaine, WA", 
                        "BN" = "Buffalo/Niagara Falls, NY", "BO" = "Boston, MA",
                        "BV" = "Brownsville, TX", "CA" = "Calais, ME", "CH" = "Chicago, IL",
                        "CL" = "Cleveland, OH", "CP" = "Champlain, NY", 
                        "CX" = "Calexico, CA", "DF" = "Dallas/Fort Worth, TX", 
                        "DE" = "Detroit, MI", "DG" = "Douglas, AZ", "DL" = "Derby Line, VT", 
                        "DN" = "Denver, CO", "DR" = "Del Rio, TX" , "DS" = "Dunseith, ND", 
                        "DU" = "Dulles, VA", "EA" = "Eastport, ID", "EL" = "El Paso, TX", 
                        "FB" = "Fairbanks, AK", "GP" = "Grand Portage, MN", 
                        "HO" = "Houlton, ME", "HA" = "Honolulu, HI","HN"= "Houston, TX", 
                        "HS" = "Highgate Springs, VT", "IF" = "International Falls, MN", 
                        "JK" = "Jackman, ME", "JU" = "Juneau, AK","LK" = "Lukeville, AZ",
                        "LA" = "Los Angeles, CA", "LO" = "Louisville, KY ","LR" = "Laredo, TX", 
                        "LV" = "Las Vegas, NV", "MC" = "McAllen, TX", "ME" = "Memphis, TN",
                        "MI" = "Miami, FL","MP" = "Minneapolis/St. Paul, MN", 
                        "NY" = "New York, NY", "NW" = "Newark, NJ","NF" = "Norfolk, VA", 
                        "NG" = "Nogales, AZ", "NO" = "New Orleans, LA", "PA" = "Philadelphia, PA", 
                        "PB" = "Pembina, ND", "PH" = "Port Huron, MI", "PL" = "Portal, ND", 
                        "PT" = "Portland, OR", "PX" = "Phoenix, AZ", "RY" = "Raymond, MT", 
                        "SF" = "San Francisco, CA","SE" = "Seattle, WA", "SJ" = "San Juan, PR", 
                        "SL" = "San Luis, AZ", "SP" = "Saipan, MP", "SS" = "Sault Saint Marie, MI", 
                        "SU" = "Sumas, WA", "SW" = "Sweetgrass, MT", 
                        "SY" = "San Diego/San Ysidro, CA", "TP" = "Tampa, FL", "XX" = "Unknown",
                        "1"= "Region 1 - Other", "2"= "Region 2 - Other","3"= "Region 3 - Other",
                        "4"= "Region 4 - Other","5"= "Region 5 - Other","6"= "Region 6 - Other",
                        "7"= "Region 7 - Other","8"= "Region 8 - Other"),
                        
         
         # Renaming descriptions
          description = recode(description, 
                               "BAL" = "Baleen", "BAR" = "Bark", "BOC" = "Bone product",
                               "BOD" = "Dead animal", "BON" = "Bone", "BOP" = "Bone piece",
                               "CAL" = "Calipee", "CAP" = "Carapace", 
                               "CAR" = "Carving", "CAV" = "Caviar", "CLA" = "Claw", 
                               "CLO"= "Cloth", "COR" = "Coral", "CPR" = "Coral product",
                               "CUL" = "Culture of an artificially propagated plant",
                               "CUT" = "Cutting", 
                               "DEA" = "Dead Specimen (died during shipment)",
                               "DER" = "Derivative", "DPL" = "Dried plant",
                               "EAR" = "Ear", "EGG" = "Egg (dead or blown)", 
                               "EGL" = "Egg (live)", "ESH" = "Eggshell",
                               "EXT" = "Extract", "FEA"= "Feather", "FIB" = "Fiber",
                               "FIG" = "Fingerling", "FIN" = "Fin", "FLO" = "Flower", 
                               "FOO" = "Foot", "GAB" = "Gall bladder", "GAL" = "Gall", 
                               "GAR" = "Garment", "GEN" = "Genitalia", 
                               "GRS" = "Graft rootstock", "HAI" = "Hair", 
                               "HAP" = "Hair product" , "HOC" = "Horn carving",
                               "HOP" = "Horn piece", "HOR" = "Horn", 
                               "IJW" = "Ivory jewelry", "IVC" ="Ivory carving",
                               "IVP" = "Ivory piece", "JWL" = "Jewelry", 
                               "KEY" = "Ivory piano key", "LPL" = "Leather product (large)", 
                               "LPS" = "Leather product (small)","LEG" = "Frog leg", 
                               "LIV" = "Live specimen", "LVS" = "Leaves", "MEA" = "Meat",
                               "MED" = "Medicinal part or product", "MUS" = "Musk",
                               "NES" = "Nest", "OIL" = "Oil", 
                               "PIV" = "Piano with ivory keys", 
                               "PLA" = "Plate of fur skins", "PLY" = "Plywood", 
                               "POW" = "Powder", "ROC" = "Coral rock (live rock)", 
                               "RUG" = "Rug", "SAW" = "Sawn wood", "SCA" = "Scale", 
                               "SEE" = "Seed", "SHE" = "Shell","SPR" = "Shell product",
                               "SHO" = "Shoe", "SID" = "Side", "SKE" = "Skeleton",
                               "SKI" = "Skin", "SKP" = "Skin piece", "SKU" = "Skull", 
                               "SOU" = "Soup", "SPE" = "Specimen (scientific or museum)",
                               "STE" = "Stem (plant)", "SWI" = "Swim bladder",
                               "TAI" = "Tail", "TEE" = "Teeth", "TIM" = "Timber", 
                               "TRI" = "Trim", "TRO" = "Trophy", "TUS" = "Tusk", 
                               "UNS" = "Unspecified", "VEN" = "Veneer", "WAX" = "Wax", 
                               "WNG" = "Wing", "WPR" = "Wood product"),
   
         
          # Renaming purposes
          purpose = recode(purpose, 
                             "B" = "Breeding in captivity/artif. propag.", 
                             "E" = "Educational", 
                             "G" = "Botanical Gardens",
                             "H" = "Hunting trophies", 
                             "M" = "Biomedical research",
                             "P" = "Personal", 
                             "Q" = "Circuses/traveling exhibitions",
                             "S" = "Scientific", 
                             "T" = "Commercial", 
                             "Y" = "Reintrod./introd. into the wild",
                             "Z" = "Zoos",
                             "L" = "Others",
                             "non-standard value" = "Others"),
         
                   # Renaming sources
          source = recode(source, 
                                   "W" = "Specimens taken from the wild", 
                                   "C" = "Animals bred in captivity", 
                                   "R" = "Specimens orig. from a ranching op.",
                                   "F" = "Animals born in captivity/not captive-bred",
                                   "U" = "Source unknown", 
                                   "D" = "commercially bred or propagated in CITES",
                                   "I" = "Confiscated or seized specimens", 
                                   "A" = "Plants art. propagated, parts and derivates",
                                   "P" = "Others", 
                                   "non-standard value" = "Others"))

# To factor
data$country_imp_exp<- as.factor(data$country_imp_exp)
data$country_origin<-as.factor(data$country_origin)

```
```{r eval=TRUE, echo=TRUE, warning=FALSE, include=TRUE}
### Defining functions

```
