# Trends over time
Let's explore how some trends change over time. 

## Shipments
**Let's summarized the number of unique shipments per month.**

```{r eval=TRUE, echo=TRUE, warning=FALSE, include=TRUE}
unique_shipments_year <-data %>% 
  group_by(year= year(shipment_date),month = month(shipment_date)) %>%
  summarise(unique_shipments = n_distinct(control_number)) 

unique_shipments_year$date <- as.yearmon(paste(unique_shipments_year$year, 
                                          unique_shipments_year$month), "%Y %m")

ggplot(unique_shipments_year, aes(x = date, y = unique_shipments)) + 
  geom_line(color='steelblue') + scale_x_continuous()
```

**Let's see the pattern depending on the week of the year**

```{r eval=TRUE, echo=TRUE, warning=FALSE, include=TRUE}
unique_shipments_week <-data %>% 
  group_by(year= year(shipment_date), week=week(shipment_date)) %>%
  summarise(unique_shipments = n_distinct(control_number)) 

ggplot(unique_shipments_week, aes(x = week, y = unique_shipments, color=factor(year))) + 
  geom_line() 

rm(unique_shipments_year)
rm(unique_shipments_week)
```

## Import (US dollars)
**It's important to take into account that since year 2008, this variable has a lot of missing values (check the 2.1. section). So, we can just calculate the total import for the period between 2000-2007, and also for the year 2012.**

### Measured by disposition year
```{r eval=TRUE, echo=TRUE, warning=FALSE, include=TRUE,out.width='90%'}
value_wildlifeproduct_dispyear <-data %>% 
  filter(!is.na(value))  %>%
  filter(!disposition_year %in% (c(2008, 2009, 2010, 2011, 2013, 2014))) %>%
  group_by(disposition_year) %>%
  summarise(value = sum(as.numeric(value))) 

DT::datatable(value_wildlifeproduct_dispyear) %>%
  formatCurrency('value',currency = "", interval = 3, mark = ",")

ggplot(value_wildlifeproduct_dispyear, aes(x = disposition_year, y=value)) + 
  geom_bar(stat = "identity", width=0.4, position = position_dodge(width=0.5), 
           fill="blue") +
  geom_text(aes(label = scales::comma(value)), vjust=-1, size=2) +
  scale_y_continuous(labels = scales::dollar_format(prefix="$")) +
  labs(y = "Dollars",  x="Disposition Year")  +
  theme(axis.text.x = element_text(hjust=1, angle=60)) 

rm(value_wildlifeproduct_dispyear)
```

### Measured by shipment year
```{r eval=TRUE, echo=TRUE, warning=FALSE, include=TRUE, out.width='90%' }
value_wildlifeproduct_shipyear <-data %>% 
  filter(!is.na(value))  %>%
  filter(!is.na(shipment_year))  %>%
  filter(!shipment_year %in% (c(2008, 2009, 2010, 2011, 2013, 2014))) %>%
  group_by(shipment_year) %>%
  summarise(value = sum(as.numeric(value)))

DT::datatable(value_wildlifeproduct_shipyear) %>%
  formatCurrency('value',currency = "", interval = 3, mark = ",")

ggplot(value_wildlifeproduct_shipyear, aes(x = shipment_year, y=value)) + 
  geom_bar(stat = "identity", width=0.4, position = position_dodge(width=0.5), 
           fill="blue") +
  geom_text(aes(label = scales::comma(value)), vjust=-1, size=2) +
  scale_y_continuous(labels = scales::dollar_format(prefix="$")) +
  labs(y = "Dollars",  x="Shipment Year") +
  theme(axis.text.x = element_text(hjust=1, angle=60)) 

rm(value_wildlifeproduct_shipyear)
```

### By taxa and shipment year
```{r eval=TRUE, echo=TRUE, warning=FALSE, include=TRUE, out.width='90%'}
value_bytaxayear<- data %>% group_by(taxa, shipment_year) %>%
  filter(!is.na(value))  %>%
  summarise(value=sum(as.numeric(value))) 

DT::datatable(value_bytaxayear) %>%
  formatCurrency('value',currency = "", interval = 3, mark = ",")
```

**Here you have a dynamic graph, where you can choose the specific taxa elements you want to display**
```{r eval=TRUE, echo=TRUE, warning=FALSE, include=TRUE, out.width='90%'}
value_bytaxayear %>%  
  plot_ly(x = ~ shipment_year) %>% 
  add_lines(y = ~ value, 
            color = ~ taxa,
             visible="legendonly")

rm(value_bytaxayear)
```
### By description and shipment year
```{r eval=TRUE, echo=TRUE, warning=FALSE, include=TRUE, out.width='90%'}
value_bydescriptionyear<- data %>% group_by(description, shipment_year) %>%
  filter(!is.na(value))  %>%
  summarise(value=sum(as.numeric(value))) 

DT::datatable(value_bydescriptionyear)  %>%
  formatCurrency('value',currency = "", interval = 3, mark = ",")

```

**Here you have a dynamic graph, where you can choose the specific descriptions you want to display**
```{r eval=TRUE, echo=TRUE, warning=FALSE, include=TRUE, out.width='90%'}
value_bydescriptionyear %>%  
  plot_ly(x = ~ shipment_year) %>% 
  add_lines(y = ~ value, 
            color = ~ description,
             visible="legendonly")

rm(value_bydescriptionyear)
```


## Quantity (number of items and kg)
**Let's summarized the quantity brought per year**. We'll only include in the analysis those items measured in "number" and "kg" units, because they represent 99.45% of the data. 

```{r eval=TRUE, echo=TRUE, warning=FALSE, include=TRUE, out.width='90%' }
quantity_year<- data %>% filter(unit=="Number"| unit=="Kilograms") 
                                           # 5,451,832 -> 5,421,730 rows

quantity_year_summary<- quantity_year %>% 
  group_by(shipment_year, unit) %>%
  summarise(quantity = sum(quantity)) 

DT::datatable(quantity_year_summary) %>% formatCurrency('quantity',
                                         currency = "",
                                         interval = 3, mark = ",")

ggplot(quantity_year_summary[quantity_year_summary$unit== "Number", ],
       aes(x = shipment_year, y=quantity)) + 
       geom_bar(stat = "identity", width=0.4, 
                position = position_dodge(width=0.5), fill="blue") +
       geom_text(aes(label = scales::comma(quantity)), vjust=-1, size=2) +
       labs(y = "Number",  x="Shipment Year") +
  theme(axis.text.x = element_text(hjust=1, angle=60)) 

ggplot(quantity_year_summary[quantity_year_summary$unit== "Kilograms", ],
       aes(x = shipment_year, y=quantity)) + 
       geom_bar(stat = "identity", width=0.4, 
                position = position_dodge(width=0.5), fill="blue") +
       geom_text(aes(label = scales::comma(quantity)), vjust=-1, size=2) +
       labs(y = "Kilograms",  x="Shipment Year") +
  theme(axis.text.x = element_text(hjust=1, angle=60)) 

rm(quantity_year, quantity_year_summary)
```
