# Trends over time
The trends analysis will focus on the number of shipments, the import (US dollars), the contents and the capacity of deter ilegal shipments (refused shipments). 

## Number of shipments
**In the 15 years analyzed, there are 2,079,637 unique shipments containing 5,451,832 segments**. We can see a clear upward trend in the number of cases from 2000 to 2006. Then, the curve reached its peak with around 162,000 unique shipments, decreased a bit, and started leveling off. 

```{r eval=TRUE, echo=TRUE, warning=FALSE, include=TRUE}
unique_shipments_yearmonth <-data %>% 
  group_by(year= year(shipment_date),month = month(shipment_date)) %>%
  summarise(unique_shipments = n_distinct(control_number)) 

unique_shipments_yearmonth$date <- as.yearmon(paste(unique_shipments_yearmonth$year, 
                                          unique_shipments_yearmonth$month), "%Y %m")

unique_shipments_year <- unique_shipments_yearmonth %>%
  ungroup() %>%
  group_by(year) %>%
  summarise(unique_shipments = sum(unique_shipments))

# Plotly by year
unique_shipments_year %>%
plot_ly(x=~year, y=~unique_shipments,type = 'scatter', mode = 'lines') %>%
    layout(title = "<b>Unique shipments per year</b>",
         xaxis= list(title= "<b>Year</b>"),
         yaxis = list(title = "<b>Total of unique shipments</b>"))


# Plotly by year and month
unique_shipments_yearmonth %>%
  plot_ly(x = ~date, y=~unique_shipments, type = 'scatter', mode = 'lines') %>%
    layout(title = "<b>Unique shipments per year and month</b>",
         xaxis= list(title= "<b>Date</b>"),
         yaxis = list(title = "<b>Total of unique shipments</b>"))


```

**Let's see the pattern depending on the week of the year**
We have excluded from the figure the 1st, 52nd and 53rd week of the year.  

```{r eval=TRUE, echo=TRUE, warning=FALSE, include=TRUE}
unique_shipments_week <-data %>% 
  group_by(year= year(shipment_date), week=week(shipment_date)) %>%
  summarise(unique_shipments = n_distinct(control_number)) %>%
  ungroup() %>%
  filter(!week %in% c(1, 52, 53)) %>%
  mutate(year=as.factor(year))

plot_ly(unique_shipments_week,x = ~week, y = ~unique_shipments) %>%
  add_lines(color=~year, colors="Dark2") %>%
  layout(title = "<b>Number of unique shipments depending on the week of the year</b>",
         xaxis= list(title= "<b>Week</b>" ,tickangle=-65),
         yaxis = list(title = "<b>Unique shipments</b>"))

rm(unique_shipments_year,unique_shipments_yearmonth,unique_shipments_week)
```

## Import (US dollars)
**It's important to take into account that since year 2008, this variable has a lot of missing values (check the 2.1. section)** So, we can just calculate the total import for the period between 2000-2007, and also for the year 2012.

It seems there has been an upward trend in the amount of spent US dollars from 2000 to 2012. 

### Measured by disposition year
```{r eval=TRUE, echo=TRUE, warning=FALSE, include=TRUE,out.width='90%'}
value_wildlifeproduct_dispyear <-data %>% 
  dplyr::filter(!is.na(value))  %>%
  dplyr::filter(!disposition_year %in% (c(2008, 2009, 2010, 2011, 2013, 2014))) %>%
  group_by(disposition_year) %>%
  summarise(value = sum(as.numeric(value))) 

value_wildlifeproduct_dispyear$disposition_year <- droplevels(value_wildlifeproduct_dispyear$disposition_year)

DT::datatable(value_wildlifeproduct_dispyear, 
              caption = htmltools::tags$caption(
                style='caption-side: bottom; text-align: center;','Table 23: ', 
                htmltools::em('Import (US dollars) by disposition year'
              ))) %>%
  formatCurrency('value', currency = "$") 
  
value_wildlifeproduct_dispyear %>% 
  plot_ly(x=~disposition_year, y=~value, 
        color=~disposition_year) %>% 
  add_bars() %>%
  layout(title = "<b>Import (US dollars) by disposition year</b>",
         xaxis= list(title= "<b>Disposition Year</b>" ,tickangle=-65),
         yaxis = list(title = "<b>US dollars</b>"))

rm(value_wildlifeproduct_dispyear)
```

### Measured by shipment year
```{r eval=TRUE, echo=TRUE, warning=FALSE, include=TRUE, out.width='90%' }
value_wildlifeproduct_shipyear <-data %>% 
  filter(!is.na(value))  %>%
  filter(!is.na(shipment_year))  %>%
  filter(!shipment_year %in% (c(2008, 2009, 2010, 2011, 2013, 2014))) %>%
  group_by(shipment_year) %>%
  summarise(value = sum(as.numeric(value)))

value_wildlifeproduct_shipyear$shipment_year <- droplevels(value_wildlifeproduct_shipyear$shipment_year)

DT::datatable(value_wildlifeproduct_shipyear, 
              caption = htmltools::tags$caption(
                style='caption-side: bottom; text-align: center;','Table 24: ', 
                htmltools::em('Import (US dollars) by shipment year'
              ))) %>%
  formatCurrency('value', currency = "$") 
  
value_wildlifeproduct_shipyear %>% 
  plot_ly(x=~shipment_year, y=~value, 
        color=~shipment_year) %>% 
  add_bars() %>%
  layout(title = "<b>Import (US dollars) by shipment year</b>",
         xaxis= list(title= "<b>Shipment Year</b>" ,tickangle=-65),
         yaxis = list(title = "<b>US dollars</b>"))

rm(value_wildlifeproduct_shipyear)
```

### By taxa and shipment year
**Items that involved mammals, shells and reptiles generate the highest amount of money throughout the years**

```{r eval=TRUE, echo=TRUE, warning=FALSE, include=TRUE, out.width='90%'}
value_bytaxayear<- data %>% 
  group_by(taxa, shipment_year) %>%
  filter(!is.na(value))  %>%
  summarise(value=sum(as.numeric(value))) %>%
  dplyr::ungroup() %>%
  filter(!shipment_year %in% (c(2008, 2009, 2010, 2011, 2013, 2014)))

DT::datatable(value_bytaxayear, filter = 'top',
              caption = htmltools::tags$caption(
                style='caption-side: bottom; text-align: center;','Table 25: ', 
                htmltools::em('Import (US dollars) by shipment year & taxa'
              ))) %>%
  formatCurrency('value', currency = "$",interval = 3, mark = ",") 

```

**Here you have a dynamic graph, where you can choose the specific taxa elements you want to display**
```{r eval=TRUE, echo=TRUE, warning=FALSE, include=TRUE, out.width='90%'}
value_bytaxayear$shipment_year<- as.numeric(as.character(value_bytaxayear$shipment_year))

value_bytaxayear <- highlight_key(value_bytaxayear)

widgets <- bscols(
  widths = c(12, 12, 12),
  filter_select("taxa","taxa", value_bytaxayear, ~taxa),
  filter_slider("Shipment Year", "Years", value_bytaxayear, ~shipment_year)
)
bscols(
  widths = c(4, 8), widgets, 
  plot_ly(value_bytaxayear, x = ~shipment_year, y = ~value) %>% 
    add_lines(color = ~taxa, colors = "Dark2") %>%
  layout(title = "<b>Import (US dollars) by shipment year & taxa</b>",
         xaxis= list(title= "<b>Shipment Year</b>" ,tickangle=-65),
         yaxis = list(title = "<b>US dollars</b>"))
)

rm(widgets,value_bytaxayear)

```
### By description and shipment year
```{r eval=TRUE, echo=TRUE, warning=FALSE, include=TRUE, out.width='90%'}
value_bydescriptionyear<- data %>% group_by(description, shipment_year) %>%
  filter(!is.na(value))  %>%
  summarise(value=sum(as.numeric(value)))  %>%
  dplyr::ungroup() %>%
  filter(!shipment_year %in% (c(2008, 2009, 2010, 2011, 2013, 2014)))

DT::datatable(value_bydescriptionyear, filter = 'top',
              caption = htmltools::tags$caption(
                style='caption-side: bottom; text-align: center;','Table 26: ', 
                htmltools::em('Import (US dollars) by shipment year & description'
              ))) %>%
  formatCurrency('value', currency = "$",interval = 3, mark = ",") 

```

**Here you have a dynamic graph, where you can choose the specific descriptions you want to display**
```{r eval=TRUE, echo=TRUE, warning=FALSE, include=TRUE, out.width='90%'}
value_bydescriptionyear$shipment_year<- as.numeric(as.character(value_bydescriptionyear$shipment_year))

value_bydescriptionyear <- highlight_key(value_bydescriptionyear)

widgets <- bscols(
  widths = c(12, 12, 12),
  filter_select("description","description", value_bydescriptionyear, ~description),
  filter_slider("Shipment Year", "Years", value_bydescriptionyear, ~shipment_year)
)

bscols(
  widths = c(4, 8), widgets, 
  plot_ly(value_bydescriptionyear, x = ~shipment_year, y = ~value) %>% 
    add_lines(color = ~description, colors = "Dark2") %>%
  layout(title = "<b>Import (US dollars) by shipment year & description</b>",
         xaxis= list(title= "<b>Shipment Year</b>" ,tickangle=-65),
         yaxis = list(title = "<b>US dollars</b>"))
)

rm(widgets, value_bydescriptionyear)
```

### By name (generic + specific) and shipment year
```{r eval=TRUE, echo=TRUE, warning=FALSE, include=TRUE, out.width='90%'}
value_byname<- data %>% 
  group_by(specific_name, shipment_year) %>%
  filter(!is.na(value))  %>%
  summarise(value=sum(as.numeric(value)))  %>%
  dplyr::ungroup() %>%
  filter(!shipment_year %in% (c(2008, 2009, 2010, 2011, 2013, 2014)))

DT::datatable(value_byname, filter = 'top',
              caption = htmltools::tags$caption(
                style='caption-side: bottom; text-align: center;','Table 27: ', 
                htmltools::em('Import (US dollars) by shipment year & specific name'
              ))) %>%
  formatCurrency('value', currency = "$",interval = 3, mark = ",") 

```
**Here you have a dynamic graph, where you can choose the specific names you want to display**. We have selected the top 100 based on the total spent US dollars. 

```{r eval=TRUE, echo=TRUE, warning=FALSE, include=TRUE, out.width='90%'}
total_byname<- data %>% 
  group_by(specific_name) %>%
  filter(!is.na(value))%>%
  summarise(value=sum(as.numeric(value))) %>%
  arrange(desc(value)) %>%
  top_n(100)

value_byname<- value_byname %>%
  mutate(shipment_year=as.numeric(as.character(shipment_year))) %>%
  dplyr::filter(specific_name %in% total_byname$specific_name)

value_byname <- highlight_key(value_byname)

widgets <- bscols(
  widths = c(12, 12, 12),
  filter_select("specific name","specific name", value_byname, ~specific_name),
  filter_slider("Shipment Year", "Years", value_byname, ~shipment_year)
)

bscols(
  widths = c(4, 8), widgets, 
  plot_ly(value_byname, x = ~shipment_year, y = ~value) %>% 
    add_lines(color = ~specific_name, colors = "Dark2") %>%
  layout(title = "<b>Import (US dollars) by shipment year & specific name</b>",
         xaxis= list(title= "<b>Shipment Year</b>" ,tickangle=-65),
         yaxis = list(title = "<b>US dollars</b>"))
)

rm(widgets, total_byname, value_byname)

```

## Content (number of items and kg)
**Let's summarized the quantity brought per year**. We'll only include in the analysis those items measured in "number" and "kg" units, because they represent 99.45% of the data. 

- It includes around 11,5 billions and 1,1 billion kg of wildlife items. 

```{r eval=TRUE, echo=TRUE, warning=FALSE, include=TRUE, out.width='90%' }
quantity_year<- data %>% filter(unit=="Number"| unit=="Kilograms") 
                                 
quantity_year_summary<- quantity_year %>% 
  group_by(shipment_year, unit) %>%
  summarise(quantity = sum(quantity)) 

DT::datatable(quantity_year_summary, filter = 'top',
              caption = htmltools::tags$caption(
                style='caption-side: bottom; text-align: center;','Table 28: ', 
                htmltools::em('Quantity by shipment year'
              ))) %>%
  formatCurrency('quantity', currency = "",interval = 3, mark = ",")

quantity_year_summary[quantity_year_summary$unit== "Number", ] %>% 
  plot_ly(x=~shipment_year, y=~quantity) %>% 
  add_bars() %>%
  layout(title = "<b>Quantity (number of items) by shipment year</b>",
         xaxis= list(title= "<b>Shipment Year</b>" ,tickangle=-65),
         yaxis = list(title = "<b>Quantity (number of items)</b>"))

quantity_year_summary[quantity_year_summary$unit== "Kilograms", ] %>% 
  plot_ly(x=~shipment_year, y=~quantity) %>% 
  add_bars() %>%
  layout(title = "<b>Quantity (Kilograms) by shipment year</b>",
         xaxis= list(title= "<b>Shipment Year</b>" ,tickangle=-65),
         yaxis = list(title = "<b>Quantity (Kilograms)</b>"))

rm(quantity_year, quantity_year_summary)
```

## Refused shipments
As we found in section 4.1.4., just 1.73% of shipments were refused. 
This would be the trend over time 

```{r eval=TRUE, echo=TRUE, warning=FALSE, include=TRUE, out.width='90%' }
refused_shipments<- data %>% 
  filter(!is.na(action))  %>%
  group_by(shipment_year, action) %>%
  summarise(total = n()) %>%
  mutate(percentage = total/sum(total)) %>%
  dplyr::filter(action=="Refused")

DT::datatable(refused_shipments, 
              caption = htmltools::tags$caption(
                style='caption-side: bottom; text-align: center;','Table 29: ', 
                htmltools::em('Refused shipments by shipment year'
              ))) %>%
  formatRound('total',1) %>%
  formatPercentage('percentage',2)

refused_shipments%>% mutate(percentage=percentage*100) %>%
  plot_ly(x=~shipment_year, y=~percentage) %>% 
  add_bars() %>%
  layout(title = "<b>Refused shipments by shipment year</b>",
         xaxis= list(title= "<b>Shipment Year</b>" ,tickangle=-65),
         yaxis = list(title = "<b>Percentage of refused shipments</b>"))

rm(refused_shipments)
```


