---
title: "Faunalytics Wildlife"
date: "10/03/2020"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 4
    toc_float: 
      collapsed: false
      smooth_scroll: true
---

<html>
  <head>
  <style type="text/css">
    * {
      text-align: justify;
    }
    
    h2 {
      color: MidnightBlue;    
      font-weight: bold;
    }
    
    h3 {
      color: MidnightBlue;    
    }  
    
    h4 {
       color: SteelBlue;    
     
    }
    
  </style>
  </head>
  <body>
  
## INTRODUCTION ##
**This report is intended to explore some ideas about LEMI dataset** accessed with the lemis R package. 
LEMIS wildlife trade data trends from 2000 through 2014. You can read more about this dataset [Here](https://www.nature.com/articles/s41597-020-0354-5)

```{r eval=TRUE, echo=TRUE, warning=FALSE, include=TRUE} 
#Load Libraries: p_load can install, load,  and update packages
pacman::p_load(rstudioapi,dplyr, ggplot2, lubridate, devtools, tidyr,magrittr, 
               lemis,zoo,gdata, knitr, DT, plotly)

# Set working directory 
path <- getwd()
knitr::opts_knit$set(root.dir = normalizePath(path.expand(path), 
                                              winslash = ("/"), 
                                              mustWork = TRUE))

# Reading dataset 
data<-readRDS("../data/data.rds")

# Transform some variables to factor/numeric/datetime
data %<>% mutate_at(c("control_number", "species_code", "taxa", "class", "genus",
                      "species", "subspecies", "specific_name", "generic_name",
                      "description", "country_origin", "country_imp_exp", 
                      "purpose","source", "action", "disposition", 
                      "disposition_year", "shipment_year", "import_export",
                      "port", "us_co", "foreign_co", "unit"), as.factor)

# Remove scientific notation
options(scipen=999)
rm(path)

```
### A. THINGS TO START LOOKING AT/THINKING ABOUT ###

#### 1) Excluding plants and microorganisms based on the kingdom level ####

**We classified 99.23 % of data by kingdoms using the class, taxa, genus, generic name, species and description columns**. This way,  we excluded those elements belonging to fungi, plantae, bacteria, chromista and unspecified kingdoms. 

**Based just on the kingdom designation, we were able to classify 92.67% of data**

* **Bacterias**: Gammaproteobacteria

* **Chromista**: Phaeophyceae

* **Plantae**: Cycadopsida, Liliopsida, Magnoliopsida, Pinopsida, Polypodiopsida

* **Fungi**: Agaricomycetes, Chytridiomycetes	

* **Animalia**: Actinopterygii, Amphibia, Anthozoa, Arachnida, Ascidiacea, 
Asteroidea, Aves, Bivalvia, Branchiopoda, Calcarea, Cephalaspidomorphi,
Cephalopoda, Cestoda, Chilopoda, Clitellata, Cubozoa, Demospongiae, Diplopoda,
Echinoidea, Elasmobranchii, Enteropneusta, Eurotatoria,Gastropoda, Gymnolaemata,
Hexactinellida, Hexanauplia, Holocephali, Holothuroidea, Hoplonemertea, Hydrozoa,
Insecta, Leptocardii, Malacostraca, Mammalia, Maxillopoda,Merostomata, Myxini,
Ophiuroidea, Ostracoda, Pilidiophora, Polychaeta, Polyplacophora, Pycnogonida,
Reptilia, Sarcopterygii, Scaphopoda, Scyphozoa, Secernentea, Sipunculidea,
Tentaculata, Thaliacea, Trematoda

```{r eval=TRUE, echo=TRUE, warning=FALSE, include=TRUE} 
# Classifying classes by kingdom 
classes<-data %>% drop_na(class) %>% group_by(class) %>% summarise(N=n())

animalia_kingdom <- c("Actinopterygii", "Amphibia", "Anthozoa", "Arachnida",
                      "Ascidiacea", "Asteroidea", "Aves", "Bivalvia",
                      "Branchiopoda", "Calcarea", "Cephalaspidomorphi",
                      "Cephalopoda", "Cestoda", "Chilopoda", "Clitellata",
                      "Crinoidea", "Cubozoa", "Demospongiae", "Diplopoda",
                      "Echinoidea", "Elasmobranchii", "Enteropneusta",
                      "Eurotatoria", "Gastropoda", "Gymnolaemata",
                      "Hexactinellida","Hexanauplia", "Holocephali",
                      "Holothuroidea", "Hoplonemertea", "Hydrozoa", "Insecta",
                      "Leptocardii", "Malacostraca", "Mammalia", "Maxillopoda",
                      "Merostomata", "Myxini", "Ophiuroidea", "Ostracoda",
                      "Pilidiophora", "Polychaeta", "Polyplacophora",
                      "Pycnogonida", "Reptilia", "Sarcopterygii", "Scaphopoda",
                      "Scyphozoa", "Secernentea", "Sipunculidea", "Tentaculata",
                      "Thaliacea", "Trematoda")

fungi_kingdom <-c("Agaricomycetes", "Chytridiomycetes")

plantae_kingdom<- c("Cycadopsida", "Liliopsida", "Magnoliopsida", "Pinopsida", 
                    "Polypodiopsida", "Ulvophyceae")

bacterias_kingdom <- c("Gammaproteobacteria")

chromista_kingdom <- c("Phaeophyceae")
```

```{r eval=TRUE, echo=TRUE, warning=FALSE, include=TRUE}
# Adding kingdom level based on class
data$kingdom<- ifelse(data$class %in% animalia_kingdom, "Animalia",
               ifelse(data$class %in% fungi_kingdom, "Fungi",        
               ifelse(data$class %in% plantae_kingdom, "Plantae", 
               ifelse(data$class %in% bacterias_kingdom, "Bacteria",                 
               ifelse(data$class %in% chromista_kingdom, "Chromista", "Other"                              )))))

rm(classes, animalia_kingdom, bacterias_kingdom, chromista_kingdom,
   plantae_kingdom, fungi_kingdom)

# Adding kingdom level based on genus
data$kingdom[which(data$genus=="Other live inverts" | data$genus=="Animals" |
                   data$genus=="Corals"| data$genus=="Dugesia"|data$genus=="Nemertea" |
                   data$genus=="Mollusca" | data$genus=="Xenoturbella" |
                   data$genus=="Chondrichthyes" | data$genus=="Chordata" | 
                   data$genus=="Paracatenula" |  data$genus=="Porifera")] <- "Animalia"
                                                         # 403,596  --> 299,674

# Adding kingdom level based on taxa
data$kingdom[which(data$taxa=="crustacean" | data$taxa=="fish" | data$taxa=="coral"|
                   data$taxa=="shell")] <- "Animalia"   # 299,673  --> 58,883

data$kingdom[which(data$taxa=="plant")] <- "Plantae"    #  58,883   --> 58,855

# Adding kingdom level based on generic name
data$kingdom[which(data$generic_name=="OTHER INVERTEBRATE" |   
                   data$generic_name=="SHELL" |
                   data$generic_name=="CORAL" | 
                   data$generic_name=="WORM")] <- "Animalia"  # 58,855 <- 40,277

# Adding kingdom level based on species
data$kingdom[which(data$species=="freshwater fish" | 
                   data$species=="invertebrates")] <- "Animalia" # 40,277 <- 40,275

# Adding kingdom level based on description
data$kingdom[which(data$description=="BOC" | data$description=="BOD" |
                   data$description=="BON" | data$description=="BOP" |
                   data$description=="CAP" | data$description=="CAV" | 
                   data$description=="CLA" | data$description=="CLA" | 
                   data$description=="MEA" | data$description=="EGG" | 
                   data$description=="EGL" | data$description=="FEA" | 
                   data$description=="FIG" | data$description=="IVC" | 
                   data$description=="LPL" | data$description=="LPS" |
                   data$description=="PLA" | data$description=="SKE" |
                   data$description=="SKI" | data$description=="SKP" |
                   data$description=="SKU" | data$description=="TEE" |
                   data$description=="TRO")] <- "Animalia" # 40,275 <- 32,048

data$kingdom[which(data$description=="CUT" | data$description=="DPL" |
                   data$description=="LOG" | data$description=="LVS" |
                   data$description=="WPR")] <- "Plantae" 
                                                                # 32,048 <- 32, 020

# Let's exclude those elements from fungi, plantae, bacteria, chromista and unspecified kingdoms. 
data <- data %>% filter(kingdom=="Animalia") 
data <- data %>% filter(taxa!="plant") 

```

#### 2) Distinct species in the dataset? ####
**It seems that there are 8,083 different species**.  We decided to exclude some elements like "freshwater sp.", "including goldfish", "marine sp", "in trop fish &" and "sp", due to they are not specific species. 

* The top 15 represents 26.53 % of data

* The top 50 represents 41.46 % of data

```{r eval=TRUE, echo=TRUE, warning=FALSE, include=TRUE}
data_species<- data %>% filter(species!="(freshwater sp.)" & 
                          species!= "(including goldfish)" &
                          species!="(marine sp.)" & 
                          species != "sp." & species != "in trop fish &")
                            
# n_distinct(data_species$species, na.rm = TRUE) # <- 8,083

species <-data_species %>% group_by(species) %>%
  summarise(total=n(), percentage= round(n()/nrow(data) *100,2)) %>%
  drop_na(species) %>%
  arrange(desc(total)) %>% top_n(15, total) 

DT::datatable(species) %>%
  formatCurrency('total',currency = "", interval = 3, mark = ",")

ggplot(data=species, aes(x=reorder(species, -percentage), y=percentage)) + 
  geom_bar(stat="identity",fill="steelblue") +
  geom_text(aes(label=round(percentage,2)))+
  theme(axis.text.x = element_text(hjust=1, angle=60)) +
  labs(y = "Percentage",  x="species") 

rm(data_species,species)
```


#### 3) Total value of the wildlife product(s) in US dollars, measured by disposition year ####   

```{r eval=TRUE, echo=TRUE, warning=FALSE, include=TRUE,out.width='90%'}
value_wildlifeproduct_dispyear <-data %>% 
  filter(!is.na(value))  %>%
  filter(!is.na(disposition_year))  %>%
  group_by(disposition_year) %>%
  summarise(value = sum(as.numeric(value)))

DT::datatable(value_wildlifeproduct_dispyear) %>%
  formatCurrency('value',currency = "", interval = 3, mark = ",")

ggplot(value_wildlifeproduct_dispyear, aes(x = disposition_year, y=value)) + 
  geom_bar(stat = "identity", width=0.4, position = position_dodge(width=0.5), 
           fill="blue") +
  geom_text(aes(label = scales::comma(value)), vjust=-1, size=2) +
  scale_y_continuous(labels = scales::dollar_format(prefix="$")) +
  labs(y = "Dollars",  x="Disposition Year")  +
  theme(axis.text.x = element_text(hjust=1, angle=60)) 

rm(value_wildlifeproduct_dispyear)
```

#### 4) Total value of the wildlife product(s) in US dollars, measured by shipment year ####  

```{r eval=TRUE, echo=TRUE, warning=FALSE, include=TRUE, out.width='90%' }
value_wildlifeproduct_shipyear <-data %>% 
  filter(!is.na(value))  %>%
  filter(!is.na(shipment_year))  %>%
  group_by(shipment_year) %>%
  summarise(value = sum(as.numeric(value)))

DT::datatable(value_wildlifeproduct_shipyear) %>%
  formatCurrency('value',currency = "", interval = 3, mark = ",")

ggplot(value_wildlifeproduct_shipyear, aes(x = shipment_year, y=value)) + 
  geom_bar(stat = "identity", width=0.4, position = position_dodge(width=0.5), 
           fill="blue") +
  geom_text(aes(label = scales::comma(value)), vjust=-1, size=2) +
  scale_y_continuous(labels = scales::dollar_format(prefix="$")) +
  labs(y = "Dollars",  x="Shipment Year") +
  theme(axis.text.x = element_text(hjust=1, angle=60)) 

rm(value_wildlifeproduct_shipyear)
```

#### 5)  Quantity brought by year ####  

```{r eval=TRUE, echo=TRUE, warning=FALSE, include=TRUE, out.width='90%'}
quantity_year<- data %>% filter(unit=="NO" | unit=="KG") # 5,107,119 -> 5,078,227
quantity_year_summary<- quantity_year %>% group_by(shipment_year, unit) %>%
  summarise(quantity = sum(quantity)) 

# Number per year
DT::datatable(quantity_year_summary[quantity_year_summary$unit== "NO", ])  %>%
  formatCurrency('quantity',currency = "", interval = 3, mark = ",")

ggplot(quantity_year_summary[quantity_year_summary$unit== "NO", ],
       aes(x = shipment_year, y=quantity)) + 
       geom_bar(stat = "identity", width=0.4, position = position_dodge(width=0.5), 
           fill="blue") +
       geom_text(aes(label = scales::comma(quantity)), vjust=-1, size=2) +
       labs(y = "Number",  x="Shipment Year") +
  theme(axis.text.x = element_text(hjust=1, angle=60)) 

# Kilograms per year
DT::datatable(quantity_year_summary[quantity_year_summary$unit== "KG", ]) %>%
  formatCurrency('quantity',currency = "", interval = 3, mark = ",")

ggplot(quantity_year_summary[quantity_year_summary$unit== "KG", ], 
       aes(x = shipment_year, y=quantity)) + 
       geom_bar(stat = "identity", width=0.4, position = position_dodge(width=0.5), 
           fill="blue") +
       geom_text(aes(label = scales::comma(quantity)), vjust=-1, size=2) +
       labs(y = "Kilograms",  x="Shipment Year")  +
  theme(axis.text.x = element_text(hjust=1, angle=60)) 

rm(quantity_year, quantity_year_summary)
```

#### 6)  Total value of import by taxa ####  
```{r eval=TRUE, echo=TRUE, warning=FALSE, include=TRUE, out.width='90%'}
value_bytaxa<- data %>% group_by(taxa) %>%
  filter(!is.na(value))  %>%
  summarise(value=sum(as.numeric(value))) %>%
  arrange(desc(value))

DT::datatable(value_bytaxa) %>%
  formatCurrency('value',currency = "", interval = 3, mark = ",")

ggplot(value_bytaxa, aes(x = reorder(taxa, -value), y=value)) + 
       geom_bar(stat = "identity", width=0.4, position = position_dodge(width=0.5), 
           fill="blue") +
       scale_y_continuous(labels = scales::dollar_format(prefix="$")) +
       geom_text(aes(label = scales::comma(value)), vjust=-1, size=2) +
       labs(y = "Value",  x="Taxa") +
  theme(axis.text.x = element_text(hjust=1, angle=60)) 

rm(value_bytaxa)

```

#### 7) Total value of import by taxa by year ####  

```{r eval=TRUE, echo=TRUE, warning=FALSE, include=TRUE, out.width='90%'}
value_bytaxayear<- data %>% group_by(taxa, shipment_year) %>%
  filter(!is.na(value))  %>%
  summarise(value=sum(as.numeric(value))) 

value_bytaxayear01<- value_bytaxayear %>% filter(taxa %in% c("mammal", "shell", 
                                                              "fish",  "reptile",
                                                              "reptile", "insect", 
                                                             "bird"))

value_bytaxayear02<- value_bytaxayear %>% filter(taxa %in% c("amphibian", 
                                                              "coral", "crustacean" ,
                                                              "echinoderms",
                                                             "other"))

value_bytaxayear03<- value_bytaxayear %>% filter(taxa %in% c("annelid",
                                                             "spider"))
DT::datatable(value_bytaxayear) %>%
  formatCurrency('value',currency = "", interval = 3, mark = ",")
```

**Here you have different graphs with some taxa elements**

```{r eval=TRUE, echo=TRUE, warning=FALSE, include=TRUE, out.width='90%'}
ggplot(value_bytaxayear01,aes(x = shipment_year, y=value, colour=taxa, group=1)) + 
       geom_line(aes(group = taxa)) +
       scale_y_continuous(labels = scales::dollar_format(prefix="$")) +
       geom_text(aes(label = scales::comma(value)), vjust=-1, size=2) +
       labs(y = "Value",  x="Taxa") +
        theme(axis.text.x = element_text(hjust=1, angle=60)) 

ggplot(value_bytaxayear02,aes(x = shipment_year, y=value, colour=taxa, group=1)) + 
       geom_line(aes(group = taxa)) +
       scale_y_continuous(labels = scales::dollar_format(prefix="$")) +
       geom_text(aes(label = scales::comma(value)), vjust=-1, size=2) +
       labs(y = "Value",  x="Taxa") +
       theme(axis.text.x = element_text(hjust=1, angle=60)) 

ggplot(value_bytaxayear03,aes(x = shipment_year, y=value, colour=taxa, group=1)) + 
       geom_line(aes(group = taxa)) +
       scale_y_continuous(labels = scales::dollar_format(prefix="$")) +
       geom_text(aes(label = scales::comma(value)), vjust=-1, size=2) +
       labs(y = "Value",  x="Taxa") +
       theme(axis.text.x = element_text(hjust=1, angle=60)) 
```

**Here you have a dynamic graph, where you can choose the specific taxa elements you want to display**
```{r eval=TRUE, echo=TRUE, warning=FALSE, include=TRUE, out.width='90%'}
value_bytaxayear %>%  
  plot_ly(x = ~ shipment_year) %>% 
  add_lines(y = ~ value, 
            color = ~ taxa,
             visible="legendonly")

rm(value_bytaxa, value_bytaxayear, value_bytaxayear01, value_bytaxayear02, value_bytaxayear03)

```
 
#### 8) Total value of import by description ####  

```{r eval=TRUE, echo=TRUE, warning=FALSE, include=TRUE, out.width='90%'}
value_bydescription<- data %>% group_by(description) %>%
  filter(!is.na(value))  %>%
  summarise(value=sum(as.numeric(value))) %>%
  arrange(desc(value))

DT::datatable(value_bydescription)  %>%
  formatCurrency('value',currency = "", interval = 3, mark = ",")

value_bydescription %>% top_n(15) %>%
ggplot(aes(x = reorder(description, -value), y=value)) + 
       geom_bar(stat = "identity", width=0.4, position = position_dodge(width=0.5), 
           fill="blue") +
       scale_y_continuous(labels = scales::dollar_format(prefix="$")) +
       geom_text(aes(label = scales::comma(value)), vjust=-1, size=2) +
       labs(y = "Value",  x="Description") +
       theme(axis.text.x = element_text(hjust=1, angle=60)) 

rm(value_bydescription)


```
#### 9) Total value of import by description by year ####  

```{r eval=TRUE, echo=TRUE, warning=FALSE, include=TRUE, out.width='90%'}
value_bydescriptionyear<- data %>% group_by(description, shipment_year) %>%
  filter(!is.na(value))  %>%
  summarise(value=sum(as.numeric(value))) 

value_bydescriptionyear01<- value_bydescriptionyear %>% 
  filter(description %in% c("SPR", "GAR",  "IVC",
                            "LPS", "SHO"))

value_bydescriptionyear02<- value_bydescriptionyear %>% 
  filter(description %in% c("LIV", "MEA", "SKI",
                            "BOD", "JWL"))

value_bydescriptionyear03<- value_bydescriptionyear %>% 
  filter(description %in% c("TRI", "LPL", "IVP",
                            "SPE", "SHE"))

DT::datatable(value_bydescriptionyear)  %>%
  formatCurrency('value',currency = "", interval = 3, mark = ",")

```

**Here you have different graphs with some descriptions** 
```{r eval=TRUE, echo=TRUE, warning=FALSE, include=TRUE, out.width='90%'}
ggplot(value_bydescriptionyear01,aes(x = shipment_year, y=value, 
                                     colour=description, group=1)) + 
       geom_line(aes(group = description)) +
       scale_y_continuous(labels = scales::dollar_format(prefix="$")) +
       geom_text(aes(label = scales::comma(value)), vjust=-1, size=2) +
       labs(y = "Value",  x="Description") +
       theme(axis.text.x = element_text(hjust=1, angle=60)) 

ggplot(value_bydescriptionyear02,aes(x = shipment_year, y=value, 
                                     colour=description, group=1)) + 
       geom_line(aes(group = description)) +
       scale_y_continuous(labels = scales::dollar_format(prefix="$")) +
       geom_text(aes(label = scales::comma(value)), vjust=-1, size=2) +
       labs(y = "Value",  x="Description") +
       theme(axis.text.x = element_text(hjust=1, angle=60)) 

ggplot(value_bydescriptionyear03,aes(x = shipment_year, y=value, 
                                     colour=description, group=1)) + 
       geom_line(aes(group = description)) +
       scale_y_continuous(labels = scales::dollar_format(prefix="$")) +
       geom_text(aes(label = scales::comma(value)), vjust=-1, size=2) +
       labs(y = "Value",  x="Description") +
       theme(axis.text.x = element_text(hjust=1, angle=60)) 
```

**Here you have a dynamic graph, where you can choose the specific descriptions you want to display**
```{r eval=TRUE, echo=TRUE, warning=FALSE, include=TRUE, out.width='90%'}
value_bydescriptionyear %>%  
  plot_ly(x = ~ shipment_year) %>% 
  add_lines(y = ~ value, 
            color = ~ description,
             visible="legendonly")

rm(value_bydescriptionyear, value_bydescriptionyear01, value_bydescriptionyear02, value_bydescriptionyear03)
```


### B. AVERAGE UNIT VALUES ###
#### 10) Average value of import by taxa (for "NO" units) ####  

```{r eval=TRUE, echo=TRUE, warning=FALSE, include=TRUE, out.width='90%'}
avgimport_bytaxa_no<- data %>% 
  filter(unit=="NO" & quantity != 0) %>%
  mutate(import=value/quantity) %>%
  group_by(taxa) %>%
  filter(!is.na(import))  %>%
  summarise(import_mean=round(mean(as.numeric(import)),2), 
            import_median = round(median(as.numeric(import))),2) %>%
  arrange(desc(import_median)) 

DT::datatable(avgimport_bytaxa_no)  %>%
  formatCurrency('import_median',currency = "", interval = 3, mark = ",")

ggplot(avgimport_bytaxa_no, aes(x = reorder(taxa, -import_median), y=import_median)) + 
       geom_bar(stat = "identity", width=0.4, position = position_dodge(width=0.5), 
           fill="blue") +
       scale_y_continuous(labels = scales::dollar_format(prefix="$")) +
       geom_text(aes(label = scales::comma(import_median)), vjust=-1, size=2) +
       labs(y = "Median Import",  x="Taxa") +
       theme(axis.text.x = element_text(hjust=1, angle=60)) 

rm(avgimport_bytaxa_no)
```

#### 11) Average value of import by description (for "NO" units) #### 
```{r eval=TRUE, echo=TRUE, warning=FALSE, include=TRUE, out.width='90%'}
avgimport_bydescrip_no<- data %>% 
  filter(unit=="NO" & quantity != 0) %>%
  mutate(import=value/quantity) %>%
  group_by(description) %>%
  filter(!is.na(import))  %>%
  summarise(import_mean=round(mean(as.numeric(import)),2), 
            import_median = round(median(as.numeric(import))),2) %>%
  arrange(desc(import_median))

DT::datatable(avgimport_bydescrip_no)  %>%
  formatCurrency('import_median',currency = "", interval = 3, mark = ",")

avgimport_bydescrip_no %>% top_n(15, wt=import_median) %>%
ggplot(aes(x = reorder(description, -import_median), y=import_median)) + 
       geom_bar(stat = "identity", width=0.4, position = position_dodge(width=0.5), 
           fill="blue") +
       scale_y_continuous(labels = scales::dollar_format(prefix="$")) +
       geom_text(aes(label = scales::comma(import_median)), vjust=-1, size=2) +
       labs(y = "Median Import",  x="Description") +
       theme(axis.text.x = element_text(hjust=1, angle=60)) 

rm(avgimport_bydescrip_no)

```
 
#### 12) Average value of import by taxa (for "KG" units) #### 

```{r eval=TRUE, echo=TRUE, warning=FALSE, include=TRUE, out.width='90%'}
avgimport_bytaxa_kg<- data %>% 
  filter(unit=="KG" & quantity != 0) %>%
  mutate(import=value/quantity) %>%
  group_by(taxa) %>%
  filter(!is.na(import))  %>%
  summarise(import_mean=round(mean(as.numeric(import)),2), 
            import_median = round(median(as.numeric(import))),2) %>%
  arrange(desc(import_median))

DT::datatable(avgimport_bytaxa_kg)  %>%
  formatCurrency('import_median',currency = "", interval = 3, mark = ",")

ggplot(avgimport_bytaxa_kg, aes(x = reorder(taxa, -import_median), y=import_median)) + 
       geom_bar(stat = "identity", width=0.4, position = position_dodge(width=0.5), 
           fill="blue") +
       scale_y_continuous(labels = scales::dollar_format(prefix="$")) +
       geom_text(aes(label = scales::comma(import_median)), vjust=-1, size=2) +
       labs(y = "Median Import",  x="Taxa") +
       theme(axis.text.x = element_text(hjust=1, angle=60)) 

rm(avgimport_bytaxa_kg)


```

#### 13) Average value of import by description (for "KG" units) #### 

```{r eval=TRUE, echo=TRUE, warning=FALSE, include=TRUE, out.width='90%'}
avgimport_bydescription_kg<- data %>% 
  filter(unit=="KG" & quantity != 0) %>%
  mutate(import=value/quantity) %>%
  group_by(description) %>%
  filter(!is.na(import))  %>%
  summarise(import_mean=round(mean(as.numeric(import)),2), 
            import_median = round(median(as.numeric(import))),2) %>%
  arrange(desc(import_median))

DT::datatable(avgimport_bydescription_kg)  %>%
  formatCurrency('import_median',currency = "", interval = 3, mark = ",")

avgimport_bydescription_kg %>% top_n(15, wt=import_median) %>%
ggplot(aes(x = reorder(description, -import_median), y=import_median)) + 
       geom_bar(stat = "identity", width=0.4, position = position_dodge(width=0.5), 
           fill="blue") +
       scale_y_continuous(labels = scales::dollar_format(prefix="$")) +
       geom_text(aes(label = scales::comma(import_median)), vjust=-1, size=2) +
       labs(y = "Median Import",  x="Description") +
       theme(axis.text.x = element_text(hjust=1, angle=60)) 

rm(avgimport_bytaxa_kg)


```

**Sources:** 

* https://www.iucnredlist.org/ 
* http://www.catalogueoflife.org/ 
* https://plants.usda.gov/classification.html 
